survAccuracyMeasures
=============================================

This R package estimates accuracy measures for risk prediction markers from survival data. It consists of the function `survEstMeasures` which estimates the $AUC$, $TPR(c)$, $FPR(c)$, $PPV(c)$, and $NPV(c)$ for for a specific timepoint and marker cutoff value c. Standard errors, and confidence intervals are also computed. Either analytic or bootstrap standard errors can be computed.

For more information, see references below. 


## Tutorial


```{r}
library(survAccuracyMeasures)

#simulated data for illustration
data(SimData)
head(SimData)
```


Estimate all measures, using asymptotic normality to estimate standard errors. 

```{r}
#Estimate all the measures at future time 2, with a marker cutpoint at 0. 

survEstMeasures(time =SimData$survTime, 
                event = SimData$status, 
                marker = SimData$Y, predict.time = 2, 
                 cutpoint = 0, SEmethod = "normal")
```

Only estimate the $AUC$ and $TPR(0)$. This time use bootstrapping to obtain the standard errors. 

```{r}
tmp <- survEstMeasures(time =SimData$survTime, event = SimData$status, marker = SimData$Y, predict.time = 2, measures = c("AUC", "TPR"), SEmethod = 'bootstrap', bootstraps = 50, cutpoint = 0)
tmp
```

```{r}
#access the estimates 
tmp$estimates

#and the confidence bounds
tmp$CIbounds
```

For more information see `?survEstMeasures`. 


### References
Pepe MS, Zheng Y, Jin Y. Evaluating the ROC performance of markers for future events. *Lifetime Data Analysis.* 2008, 14: 86-113.

Zheng Y, Cai T, Pepe MS, Levy, W. Time-dependent predictive values of prognostic biomarkers with failure time outcome. *JASA* 2008, 103: 362-368.














