---
output: html_document
---
survAccuracyMeasures
=============================================

This R package computes non-parametric (NP) and semi-parametric (SP) estimates of common accuracy measures for risk prediction markers from survival data. It consists of the function `survAM.estimate` which estimates the *AUC*, *TPR( c )*, *FPR( c )*, *PPV( c )*, and *NPV( c )* for for a specific prediction time and marker cutoff value c.  

 NP estimates are calculated using inverse probability weighting, while SP estimates are based on a Cox proportional hazards model. For detailed information regarding estimation methods, see references below. 

Standard errors for estimates can be obtained by bootstrapping. Asymptotic standard error calculations are also available for semi-parametric Cox estimates. Confidence intervals using a normal approximation are computed.

## Tutorial


```{r, eval=FALSE}
#install the package from github
# download the package from github
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mdbrown/survAccuracyMeasures")
```

```{r}
library(survAccuracyMeasures)
set.seed(112233)
#simulated data for illustration
data(SimData)
head(SimData)
```


Estimate all measures, using the bootstrap to estimate standard errors. First we obtain non-parametric estimates using inverse probablity weighting:

```{r}

#non-parametric estimates with bootstrap standard errrors
##in practice 'bootstraps' should be set to a larger value
survAM.estimate(time =survTime,
                event = status,
                marker = Y,
                data = SimData,
                estimation.method = "IPW",
                se.method = "bootstrap",
                predict.time = 1,
                threshold = 0.5,
                threshold.type = "marker", 
                bootstraps = 50)

survAM.estimate(time =survTime, 
                event = status, 
                marker = Y,
                data = SimData, 
                estimation.method = "IPW",    
                se.method = "asymptotic",
                predict.time = 1, 
                threshold = 0.5, 
                threshold.type = 'marker' )
```

Alternatively, we can calculate semi-parametric estimates based on a Cox proportional hazards model:


```{r}
#semi-parametric estimates assuming a cox model
##in practice 'bootstraps' should be set to a larger value
survAM.estimate(time =survTime, 
                event = status, 
                marker = Y,
                data = SimData, 
                estimation.method = "Cox",
                se.method = "bootstrap",
                predict.time = 1, 
                threshold = 0, 
                threshold.type = 'marker',
                bootstraps = 50)


## here we calculate the SE's based on the asymptotic properties of the Cox estimator.  
survAM.estimate(time =survTime, 
                event = status, 
                marker = Y,
                data = SimData, 
                estimation.method = "Cox",    
                se.method = "asymptotic",
                predict.time = 1, 
                threshold =  0.5, 
                threshold.type = 'marker')

```

The threshold can also be selected by fixing the FPR, TPR, PPV, or NPV using the `threshold.type` argument. Below we select the threshold such that the FPR is fixed to 0.05. 

```{r}
#non-parametric estimates with bootstrap standard errrors
##in practice 'bootstraps' should be set to a larger value
survAM.estimate(time =survTime, 
                event = status, 
                marker = Y,
                data = SimData, 
                estimation.method = "IPW",    
                se.method = "bootstrap",
                predict.time = 1, 
                threshold =  .05, 
                threshold.type = 'FPR', 
                bootstraps = 50)

```


## Plot ROC curves 

Information to plot an ROC curve can be accessed 

```{r}

est.ipw <- survAM.estimate(time =survTime,
                event = status,
                marker = Y,
                data = SimData,
                estimation.method = "IPW",
                se.method = "asymptotic",
                predict.time = 1,
                threshold = 0.5,
                threshold.type = "marker")
#information for roc curve 
head(est.ipw$roc) 

est.cox <- survAM.estimate(time =survTime,
                event = status,
                marker = Y,
                data = SimData,
                estimation.method = "Cox",
                se.method = "asymptotic",
                predict.time = 1,
                threshold = 0.5,
                threshold.type = "marker")

#information for roc curve 
head(est.cox$roc) 


plot(est.ipw$roc$FPR, est.ipw$roc$TPR, type = "l", 
     xlab = "FPR", ylab = "TPR", 
     col = "dodgerblue", lwd = 1.5)
lines(est.cox$roc$FPR, est.cox$roc$TPR, col = "firebrick", lwd = 1.5)
abline(0,1, lty = 2)

legend("bottomright", legend = c("IPW", "Cox"), col = c("dodgerblue", "firebrick"), lwd = c(1.5, 1.5), lty = c(1,1))

```


For more information see `?survAM.estimate`. 


### References
Liu D, Cai T, Zheng Y. Evaluating the predictive value of biomarkers with stratified case-cohort design. *Biometrics* 2012, 4: 1219-1227.

Pepe MS, Zheng Y, Jin Y. Evaluating the ROC performance of markers for future events. *Lifetime Data Analysis.* 2008, 14: 86-113.

Zheng Y, Cai T, Pepe MS, Levy, W. Time-dependent predictive values of prognostic biomarkers with failure time outcome. *JASA* 2008, 103: 362-368.














